noam_final_analyses

### 2024-09-04 noam_final_analyses.Rmd

@@

```{r read and merge secondary datasets}
wave <- read.csv(here::here("data", "algae_wave_dynamics.csv"), stringsAsFactors = F)
nutrients <- read.csv(here::here("data", "lter_nuts.csv"), stringsAsFactors = F)
temp <- read.csv(here::here("data", "temp_sum.csv"), stringsAsFactors = F)

@@

```{r all algae alpha diversity quadrat level}
alpha_diversity_quad <-
  data.frame(
    year = data$year,
    location = data$location,
    shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
    richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
  )

@@

```{r macroalgae diversity quadrat level}
alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
  )

@@

```{r all algae alpha diversity site level}
alpha_diversity_site <-
  data.frame(
    year = site_full$year,
    site_hab_year = site_full$site_hab_year,
    shannon = diversity(site_full[ , -which(names(site_full) %in% meta_cols)], "shannon"),
    richness = rowSums(site_full[ , -which(names(site_full) %in% meta_cols)] > 0)
  )

site_full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_full_richness_mod)
hist(residuals(site_full_richness_mod)) # similar slight skew to before

# striping/patterining expected because of discrete response var. otherwise fine
plot(residuals(site_full_richness_mod) ~ fitted(site_full_richness_mod)) # slightly heteroskedastic

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_full_shannon_mod <- glmmTMB(log(shannon/(hard_cover/100) +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_full_shannon_mod)
hist(residuals(site_full_shannon_mod)) # looks good

# slight patterning but generally fine
plot(residuals(site_full_shannon_mod) ~ fitted(site_full_shannon_mod))

emtrends(site_full_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ
site_full_shannon_emm <- emmip(site_full_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = shannon/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "shannon") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```{r macroalgae alpha diversity site level}
alpha_diversity_site_macro <-
  data.frame(
    year = site_macro$year,
    site_hab_year = site_macro$site_hab_year,
    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
  )

site_macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site_macro) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_macro_richness_mod)
hist(residuals(site_macro_richness_mod)) # slightly more skew

# striping/patterining expected because of discrete response var. otherwise fine
plot(residuals(site_macro_richness_mod) ~ fitted(site_macro_richness_mod))

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro,
             aes(x = year, y = richness/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10(lim = c(0.1,20)) +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_macro_shannon_mod <- glmmTMB(log(shannon/(hard_cover/100) +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_macro_shannon_mod)
hist(residuals(site_macro_shannon_mod)) # looks good

# slight patterning but generally fine
plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod))

(site_base_shannon_plot <- 
  ggplot() +
 # geom_point(data = alpha_diversity_site,
 #            aes(x = year, y = shannon/(hard_cover/100), colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_shannon_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_shannon_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
#  scale_y_log10() +
    labs(x = "Year", y = "shannon") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

>>>>

data[,full_species_cols] <- data[,full_species_cols]/data$hard_cover # transform cover to proportion of hardcover in quadrat

macroalgae[,macro_species_cols] <- macroalgae[,macro_species_cols]/macroalgae$hard_cover # repeat for macro dataframe

@@

```{r create adiv dataframes}
alpha_diversity_quad <-
  data.frame(
    year = data$year,
    location = data$location,
    shannon = diversity(data[ , -which(names(data) %in% meta_cols)], "shannon"),
    richness = rowSums(data[ , -which(names(data) %in% meta_cols)] > 0)
  )

alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
  )

alpha_diversity_site <-
  data.frame(
    year = site_full$year,
    site_hab_year = site_full$site_hab_year,
    shannon = diversity(site_full[ , -which(names(site_full) %in% meta_cols)], "shannon"),
    richness = rowSums(site_full[ , -which(names(site_full) %in% meta_cols)] > 0)
  )

alpha_diversity_site_macro <-
  data.frame(
    year = site_macro$year,
    site_hab_year = site_macro$site_hab_year,
    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
  )

@@

```{r read and merge secondary datasets, fig.height=15, fig.width=15}
wave <- read.csv(here::here("data", "algae_wave_dynamics.csv"), stringsAsFactors = F)
wave <- wave[,-which(names(wave) %in% "X")] # remove rownum columns
wave$site_hab_year <- paste0(wave$site,"_", wave$habitat,"_", wave$year) # for merge

nutrients <- read.csv(here::here("data", "lter_nuts.csv"), stringsAsFactors = F)
nutrients$site_hab_year <- paste0(nutrients$site,"_", nutrients$habitat,"_", nutrients$year)

temp <- read.csv(here::here("data", "temp_sum.csv"), stringsAsFactors = F)
temp$site_hab_year <- paste0(temp$site,"_", temp$habitat,"_", temp$year)

herb <- read.csv(here::here("data", "herbivore_biomass_and_diversity_site.csv"), stringsAsFactors = F)
# create column for total herbivore biomass
herb$herbivore_biomass <- rowSums(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))])
herb<-herb %>%  # rename columns for merge
  rename(
    year = Year,
    site = Site,
    habitat = Habitat,
    herbivore_richness = richness,
    herbivore_shannon = shannon,
    herbivore_evenness = evenness
    )
# fix site names and habitat names
herb <-
  herb %>% mutate(site = case_when(site == 1 ~ "lter_1",
                              site == 2 ~ "lter_2",
                              site == 3 ~ "lter_3",
                              site == 4 ~ "lter_4",
                              site == 5 ~ "lter_5",
                              site == 6 ~ "lter_6"))
herb <-
  herb %>% mutate(habitat = case_when(habitat == "Fringing" ~ "fringing",
                                      habitat == "Backreef" ~ "backreef",
                                      habitat == "Outer 10" ~ "outer_10")) # there is no outer 17 for herb data
herb$site_hab_year <- paste0(herb$site,"_", herb$habitat,"_", herb$year)
herb <- subset(herb, year != 2006)

### Merge with alpha diversity datasets

alpha_diversity_site <- full_join(alpha_diversity_site, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_site <- full_join(alpha_diversity_site, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site <- full_join(alpha_diversity_site,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site <- full_join(alpha_diversity_site, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_site, ~ all_algae_cover_hard + macroalgae_cover_hard + richness + shannon + coral_cover_hard + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_site_macro <- full_join(alpha_diversity_site_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_site_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_quad <- full_join(alpha_diversity_quad, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_quad <- full_join(alpha_diversity_quad, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad <- full_join(alpha_diversity_quad,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad <- full_join(alpha_diversity_quad, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, nutrients, by = c("year", "site_hab_year", "site", "habitat"))
alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, temp, by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro,herb[,-which(names(herb) %in% colnames(herb[,grep("Acanthurus.nigricans", colnames(herb)):grep("Leptoscarus.vaigiensis", colnames(herb))]) )], by = c("year", "site_hab_year", "site", "habitat"))

alpha_diversity_quad_macro <- full_join(alpha_diversity_quad_macro, wave, by = c("year", "island_side", "site_hab_year", "site", "habitat"))
pairs(data = alpha_diversity_quad_macro, ~ richness + shannon + wave_power_adcp + mean_n + n_cv + mean_temp_c + mean_temp_range + herbivore_biomass + herbivore_richness + herbivore_shannon + herbivore_evenness)

```{r all algae alpha diversity site level}

site_full_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site) # need to include offset because of the way richness is calculated
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
# This warning occurs when the optimizer visits a region of parameter space that is invalid. It is not a problem as long as the optimizer has left that region of parameter space upon convergence, which is indicated by an absence of the model convergence warnings described above.
summary(site_full_richness_mod) # low variance in random effect
Anova(site_full_richness_mod) 
#                 Chisq Df Pr(>Chisq)    
# year           1.5505  1     0.2131    
# habitat      106.8578  3     <2e-16 ***
# year:habitat   5.3674  3     0.1468    
# habitat more important than time when time is numeric
# check_mod
hist(residuals(site_full_richness_mod)) # slight skew

# striping/patterining expected because of discrete response var.
plot(residuals(site_full_richness_mod) ~ fitted(site_full_richness_mod)) # slightly heteroskedastic

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_full_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_full_shannon_mod <- glmmTMB(log(shannon + 1) ~ year*habitat + (1|site), 
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed. No offset because calculation of shannon already incorporated differences in %hard
summary(site_full_shannon_mod) # low variance again
Anova(site_full_shannon_mod)
# Response: log(shannon + 1)
#                Chisq Df Pr(>Chisq)    
# year          0.0133  1   0.908348    
# habitat      97.6274  3  < 2.2e-16 ***
# year:habitat 14.0548  3   0.002831 **

hist(residuals(site_full_shannon_mod)) # looks good
plot(residuals(site_full_shannon_mod) ~ fitted(site_full_shannon_mod)) # looks good here too

emtrends(site_full_shannon_mod, pairwise ~  habitat, var = "year") # backreef and fringing differ, backreef different fro outer17
site_full_shannon_emm <- emmip(site_full_shannon_mod,habitat ~ year,
                        at = list(year = seq(2007,2023, 1)), plotit = F, CIs = T)

(site_base_shannon_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site,
             aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_full_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
  geom_ribbon(data = site_full_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
  geom_segment(aes(y = 3, yend = 3, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```{r macroalgae alpha diversity site level}

site_macro_richness_mod <- glmmTMB(richness ~ year*habitat + (1|site) + offset(log(hard_cover)), family = poisson,
                             data = alpha_diversity_site_macro) 
# Warning in (function (start, objective, gradient = NULL, hessian = NULL,  : NA/NaN function evaluation
summary(site_macro_richness_mod) # low variance again
Anova(site_macro_richness_mod)
#                Chisq Df Pr(>Chisq)    
# year          1.1856  1    0.27621    
# habitat      87.7765  3    < 2e-16 ***
# year:habitat  7.0855  3    0.06922 .  
hist(residuals(site_macro_richness_mod)) # slightly more skew
plot(residuals(site_macro_richness_mod) ~ fitted(site_macro_richness_mod)) # similar patterns

(site_base_richness_plot <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro,
             aes(x = year, y = richness, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_richness_emm, aes(x = year, y = exp(yvar), colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_richness_emm, aes(x = year, ymin = exp(LCL), ymax = exp(UCL), fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  geom_segment(aes(y = 18, yend = 18, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_y_log10() +
    labs(x = "Year", y = "Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

site_macro_shannon_mod <- glmmTMB(log(shannon +1 ) ~ year*habitat + (1|site),
                             data = alpha_diversity_site) 
# sites with shannon div = 0 so log+1 transformed
summary(site_macro_shannon_mod)
Anova(site_macro_shannon_mod)
#                Chisq Df Pr(>Chisq)    
# year          0.0133  1   0.908348    
# habitat      97.6274  3  < 2.2e-16 ***
# year:habitat 14.0548  3   0.002831 ** 
hist(residuals(site_macro_shannon_mod)) # looks good
plot(residuals(site_macro_shannon_mod) ~ fitted(site_macro_shannon_mod)) # generally fine

(site_base_shannon_plot <- 
  ggplot() +
 geom_point(data = alpha_diversity_site,
            aes(x = year, y = shannon, colour = habitat), size = 2, alpha = 0.5) + 
  geom_line(data = site_macro_shannon_emm, aes(x = year, y = yvar, colour = habitat), size = 2) +
  geom_ribbon(data = site_macro_shannon_emm, aes(x = year, ymin = LCL, ymax = UCL, fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  geom_segment(aes(y = 2.5, yend = 2.5, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
  scale_fill_manual(values = habitat_colours) +
  scale_y_log10() +
    labs(x = "Year", y = "Shannon Index (H')") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank(),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) ) # add box around legend

```
```{r stepwise models site level}

## FRINGING - exclude wave power
fringing_all_site <- na.omit(subset(alpha_diversity_site, habitat == "fringing")) # only use rows that contain all factors

fringing_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass) # universal model has convergence issues
fringing_all_richness_dredge <- dredge(fringing_all_richness_mod_universal) # most models have convergence issues
model.sel(fringing_all_richness_dredge) # top model just includes year
fringing_all_richness_top_mod <- glmmTMB(richness ~ year + offset(log(hard_cover)) + (1|site), family = poisson, data = fringing_all_site, na.action = na.pass)
summary(fringing_all_richness_top_mod)
hist(residuals(fringing_all_richness_top_mod)) # slight skew but not bad
plot(residuals(fringing_all_richness_top_mod) ~ fitted(fringing_all_richness_top_mod)) # looks fine

fringing_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_dredge <- dredge(fringing_all_shannon_mod_universal) 
model.sel(fringing_all_shannon_dredge) # tie for the top model - best includes year and coral cover or just nitrogen
fringing_all_shannon_top_mod1 <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + (1|site), data = fringing_all_site, na.action = na.pass)
fringing_all_shannon_top_mod2 <- glmmTMB(log(shannon + 1) ~ mean_n + mean_temp_c + (1|site), data = fringing_all_site, na.action = na.pass)
hist(residuals(fringing_all_shannon_top_mod1)) # slightly bimodal
hist(residuals(fringing_all_shannon_top_mod2)) # also uniform - neither are great
plot(residuals(fringing_all_shannon_top_mod1) ~ fitted(fringing_all_shannon_top_mod1)) # slight patterning
plot(residuals(fringing_all_shannon_top_mod2) ~ fitted(fringing_all_shannon_top_mod2)) # similar here

## Backreef - includes all 
backreef_all_site <- na.omit(subset(alpha_diversity_site, habitat == "backreef")) # only use rows that contain all factors

backreef_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass) # universal model has convergence issues
backreef_all_richness_dredge <- dredge(backreef_all_richness_mod_universal) # most models have convergence issues
model.sel(backreef_all_richness_dredge) # top model includes year and coral 
backreef_all_richness_top_mod <- glmmTMB(richness ~ year + coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = backreef_all_site, na.action = na.pass)
summary(backreef_all_richness_top_mod)
hist(residuals(backreef_all_richness_top_mod)) # slight skew but not bad
plot(residuals(backreef_all_richness_top_mod) ~ fitted(backreef_all_richness_top_mod)) # looks fine

backreef_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp + mean_n + mean_temp_c +  herbivore_evenness + (1|site), data = backreef_all_site, na.action = na.pass)
backreef_all_shannon_dredge <- dredge(backreef_all_shannon_mod_universal) 
model.sel(backreef_all_shannon_dredge) #  best includes just coral cover 
backreef_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  coral_cover_hard + (1|site), data = backreef_all_site, na.action = na.pass)
hist(residuals(backreef_all_shannon_top_mod)) # pretty good
plot(residuals(backreef_all_shannon_top_mod) ~ fitted(backreef_all_shannon_top_mod)) # looks fine

## outer10 - no nitrogen 
outer10_all_site <- subset(alpha_diversity_site, habitat == "outer_10") # only use rows that contain all factors
# remove nitrogent columns
outer10_all_site <- outer10_all_site[,-which(names(outer10_all_site) %in% c("mean_n", "n_cv"))]
# remove other na rows
outer10_all_site <- na.omit(outer10_all_site)

outer10_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass) # universal model has convergence issues
outer10_all_richness_dredge <- dredge(outer10_all_richness_mod_universal) # most models have convergence issues
model.sel(outer10_all_richness_dredge) # top model includes just intercept... next best is coral 
outer10_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer10_all_site, na.action = na.pass)
summary(outer10_all_richness_top_mod)
hist(residuals(outer10_all_richness_top_mod)) #  not bad
plot(residuals(outer10_all_richness_top_mod) ~ fitted(outer10_all_richness_top_mod)) # looks fine

outer10_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
outer10_all_shannon_dredge <- dredge(outer10_all_shannon_mod_universal) # some convergence issues
model.sel(outer10_all_shannon_dredge) #  best includes just herbivore evenness - first time there's been a decent gap in AICc
outer10_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  herbivore_evenness + (1|site), data = outer10_all_site, na.action = na.pass)
hist(residuals(outer10_all_shannon_top_mod)) # pretty good
plot(residuals(outer10_all_shannon_top_mod) ~ fitted(outer10_all_shannon_top_mod)) # looks fine

## outer17 - no nitrogen or herbivores
outer17_all_site <- subset(alpha_diversity_site, habitat == "outer_17") # only use rows that contain all factors
# remove nitrogent columns
outer17_all_site <- outer17_all_site[,-which(names(outer17_all_site) %in% c("mean_n", "n_cv", "herbivore_biomass", "herbivore_evenness", "herbivore_richness", "herbivore_shannon"))]
# remove other na rows
outer17_all_site <- na.omit(outer17_all_site)

outer17_all_richness_mod_universal <- glmmTMB(richness ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c +  offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass) # universal model has convergence issues
outer17_all_richness_dredge <- dredge(outer17_all_richness_mod_universal) # most models have convergence issues
model.sel(outer17_all_richness_dredge) # top model includes just intercept... next best is coral 
outer17_all_richness_top_mod <- glmmTMB(richness ~ coral_cover_hard + offset(log(hard_cover)) + (1|site), family = poisson, data = outer17_all_site, na.action = na.pass)
summary(outer17_all_richness_top_mod)
hist(residuals(outer17_all_richness_top_mod)) #  not bad
plot(residuals(outer17_all_richness_top_mod) ~ fitted(outer17_all_richness_top_mod)) # looks fine

outer17_all_shannon_mod_universal <- glmmTMB(log(shannon + 1) ~ year + coral_cover_hard + wave_power_adcp  + mean_temp_c + (1|site), data = outer17_all_site, na.action = na.pass)
outer17_all_shannon_dredge <- dredge(outer17_all_shannon_mod_universal) # some convergence issues
model.sel(outer17_all_shannon_dredge) #  best includes just year
outer17_all_shannon_top_mod <- glmmTMB(log(shannon + 1) ~  year + (1|site), data = outer17_all_site, na.action = na.pass)
hist(residuals(outer17_all_shannon_top_mod)) # pretty good
plot(residuals(outer17_all_shannon_top_mod) ~ fitted(outer17_all_shannon_top_mod)) # looks fine
```

```{r all algae alpha diversity quadrat level}

@@

```{r macroalgae diversity quadrat level}

@@
